Bees need larger brains to thrive in urban environments

### 2023-09-03 Manuscript/Supplementary_material.Rmd
Work in progress for the review of Biol Lett

>>>>

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=8, fig.cap="\\textbf{Figure SX}. ", fig.pos="p", out.width="100%"}

#To answer one of the issues raised by a reviewer
#We explore how the bee's diet (oligo/polylectic)
#is associated with habitat preference (1) and brain size (2)

#The data of the diet comes from the paper of Ferran
#but because is missing for some species
#I have look the diets for missing ones (~20spp)

#Load libraries
library(dplyr)
library(stringr)
library(ggplot2)
library(patchwork)
library(readr)

# (1) Lecticity ~ habitat preference ----
#First we explore this association with habitat preference
#Load preferences data
preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
#Bind rows
pref = bind_rows(preferences_usa, preferences_eu)

#Average preferences between USA and Europe
preferences = pref %>% 
group_by(Species) %>%
summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
ungroup()

#Read lecticity
lecticity = read_csv("../Data/Bees_Lecticity.csv") 
#Delete underscore from tree
lecticity$species = str_replace(lecticity$species, "_", " ")
lecticity = lecticity %>% rename(Species = species)

#Bind datasets
d = left_join(preferences, lecticity)

#Check species with lacking info
#Agapostemon virescens
#https://lopezuribelab.com/halictidae/agapostemon-virescens/#:~:text=Agapostemon%20virescens%20is%20a%20polylectic,coneflowers%2C%20asters%2C%20and%20goldenrod.
#As the other Agapostemon, they tend to visit many unrelated plant species
#but they tend to prefer Asteraceae flowers
#Polylectic
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Agapostemon virescens", "Polylectic"))
#Andrena angustior
#http://dx.doi.org/10.1016/j.biocon.2017.09.009
#Polylectic
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena angustior", "Polylectic"))
#Andrena hirticincta
#https://www.sharpeatmanguides.com/andrena-mining-bees
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena hirticincta", "Oligolectic"))
#Andrena milwaukeensis
#https://www.discoverlife.org/20/q?search=Andrena+milwaukeensis
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena milwaukeensis", "Polylectic"))
#Andrena pilipes
#https://www.wildbienen.de/eb-apili.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena pilipes", "Polylectic"))
#Andrena vicina
#https://www.discoverlife.org/20/q?search=Andrena+vicina
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena vicina", "Polylectic"))
#Andrena wilkella
##http://dx.doi.org/10.1016/j.biocon.2017.09.009
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Andrena wilkella", "Oligolectic"))

#Augochlorella aurata
#https://www.beesofcanada.com/species/augochlorella-aurata
#https://www.discoverlife.org/20/q?search=Augochlorella+aurata
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Augochlorella aurata", "Polylectic"))

#Augochloropsis metallica
#https://www.gbif.org/species/1353091
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Augochloropsis metallica", "Polylectic"))

#Bombus griseocollis
#https://doi.org/10.1002/ecy.2697
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Bombus griseocollis", "Polylectic"))

#Calliopsis andreniformis
#https://www.sharpeatmanguides.com/calliopsis-andreniformis
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Calliopsis andreniformis", "Polylectic"))

#Ceratina strenua
#https://www.beesofcanada.com/species/ceratina-strenua
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Ceratina strenua", "Polylectic"))

#Colletes thoracicus
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Colletes thoracicus", "Polylectic"))

#Osmia atriventris
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia atriventris", "Polylectic"))

#Osmia lignaria
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia lignaria", "Polylectic"))

#Osmia pumila
#https://doi.org/10.1073/pnas.1218503110
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia pumila", "Polylectic"))

#Osmia cornifrons
#http://dx.doi.org/10.1098/rsos.200225
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Osmia cornifrons", "Polylectic"))

#Lasioglossum malachurum
#https://www.wildbienen.de/eb-lmala.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Lasioglossum malachurum", "Polylectic"))

#Megachile texana
#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))

#Megachile texana
#https://entnemdept.ufl.edu/hallg/melitto/floridabees/litomegachile.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile texana", "Polylectic"))

#Megachile willughbiella
#https://www.wildbienen.de/eb-mwill.htm
d = d %>% 
mutate(Lecticity = replace(Lecticity, Species == "Megachile willughbiella", "Polylectic"))

#Natural
p1_1 = ggplot(d, aes(x = Lecticity, y = Natural)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Habitat Preference")+
xlab(NULL)+
ggtitle("Natural")+
theme_bw()

#Agricultural
p1_2 = ggplot(d, aes(x = Lecticity, y = Agricultural)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab(NULL)+
xlab("Diet")+
ggtitle("Agricultural")+
theme_bw()

#Urban
p1_3 = ggplot(d, aes(x = Lecticity, y = Urban)) +
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab(NULL)+
xlab(NULL)+
ggtitle("Urban")+
theme_bw()

panel1 = p1_1+p1_2+p1_3

# (2) Lecticity ~ brain size ----
#Load brain data
brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")

d1 = left_join(brain_weight, d) %>% 
filter(!is.na(Lecticity))

p2_1 = ggplot(d1, aes(x = Lecticity, y = residuals)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Relative brain size")+
xlab("Diet")+
theme_bw()

p2_2 = ggplot(d1, aes(x = Lecticity, y = Brain.weight)) + 
geom_violin(alpha=0.4, fill="#d7e1ee",linewidth=0)+
geom_boxplot(colour="black",fill="grey",
width = .15, 
outlier.shape = NA) +
geom_jitter(alpha=0.3,width = 0.1) +
ylab("Absolute brain size")+
xlab("Diet")+
theme_bw()

panel2 =  p2_2 + p2_1

library(cowplot)

plot_grid(plot_grid(panel1, labels = c('A'))+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)),

plot_grid(panel2, labels = c('B'))+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1)), nrow = 2)
